Data Exploration: IUCN and AquaMaps

### 2016-03-21 data_explore/data_explore.Rmd
cleaning up data explore

``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
```
<!-- # bar chart: number of spp per group
# scatter plot: area compare IUCN vs AM, including log plot and plot by species group
# rangemap: Chromis alpha -->

``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
```
<!-- percent overlap vs total area exploration -->

***

Quadrant map and quad breakdown by taxa and risk

### mongo plot time
scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
                                         transform(popn_category = reorder(popn_category, category_score)),
                                       aes(x = area_ratio, 
                                           y = sm_perc, 
                                           color = popn_category,
                                           size = lg_area_pct)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
#             fill= "green4")  + 
             fill= "#4dac26")  + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
#             fill= "greenyellow") + 
             fill= "#b8e186") + 
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
#             fill= "orange") + 
             fill= "#f1b6da") + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
#             fill= "orangered2") + 
             fill= "#d01c8b") + 
  geom_point(alpha = .6)

### Manage scales for color and size 
scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
  scale_colour_manual(values = c('LC' = 'green4', 
                                 'NT' = 'greenyellow', 
                                 'VU' = 'yellow', 
                                 'EN' = 'orange', 
                                 'CR' = 'orangered2',
                                 'DD' = 'grey30')) + 
  scale_size_continuous(limits = c(0, 100),
                        breaks = c( 0,    25,    50,    75,    100), 
                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

### here are mean labels:
scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
  annotate(geom = 'text',
           x = area_align_mean, y = 5,
           hjust = 0, vjust = 0,
           color = 'grey30', 
           size = 2,
           fontface = 'bold.italic', angle = 90,
           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
  annotate(geom = 'text',
           x = 5, y = dist_align_mean,
           hjust = 0, vjust = 0,
           color = 'grey30', 
           size = 2,
           fontface = 'bold.italic', angle = 0,
           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))

### here are mean lines:
# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
#   geom_vline(x = area_align_mean,
#              color = 'grey50', linetype = 'dashed') +
#   geom_hline(y = dist_align_mean,
#              color = 'grey50', linetype = 'dashed')

### here are individual species, identified and namean on the plot:
# scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
#   geom_text(data = spp_list_label, 
#             aes(label = sciname_label), 
#             color = 'grey30',
#             fontface = 'bold.italic',
#             size = 2,
#             hjust = .25,
#             vjust = 1.5)

scatter_arearatio_vs_overlap <- scatter_arearatio_vs_overlap +
  labs(x = bquote('Area alignment '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
#       title = 'Distribution alignment vs Extent alignment',
       color = 'Risk', size = '% ocean range')

print(scatter_arearatio_vs_overlap)

ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 17.8, height = 9, units = 'cm')

print(barchart_spp_gp_quads)
ggsave(plot = barchart_spp_gp_quads,
       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
       width = 17.8, height = 7, units = 'cm')

print(barchart_spp_risk_quads)
ggsave(plot = barchart_spp_risk_quads,
       filename = file.path(dir_fig, 'barchart_spp_risk_quads.png'),
       width = 17.8, height = 7, units = 'cm')

<!-- # Multiple species maps per quadrant -->
``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
```

***

Breakdown maps by taxa

@@

# taxa_df_long3 <- taxa_df %>%
#   dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
#   gather(dataset, n_spp, AquaMaps:Both) %>%
#   group_by(class) %>%
#   mutate(n_total = sum(n_spp, na.rm = TRUE),
#          percent = n_spp/n_total * 100)
#
# bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
#   ggtheme_plot + 
#   geom_bar(stat = 'identity', alpha = .5) +
#   scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
#   geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
#             #size = 3, 
#             angle = 90, hjust = 0, color = 'grey30') +
#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
#   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
#   labs(# title = 'Species maps by taxonomic class, Red List-assessed only',
#        x = 'Class', 
#        y = 'Percent of assessed mapped species')
# print(bar_taxa3)
# ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), 
#        width = 17.8, height = 6, units = 'cm')

@@

***

# Years for IUCN maps?

``` {r get_iucn_citation_info, echo = FALSE, warning = FALSE, eval = FALSE}

  ### Queries shapefiles and extracts column headers.  Use to help identify which
  ### species groups have no id_no field, and possibly other fields to help 
  ### differentiate.

dir_iucn_shp <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp/iucn_shp')
groups_list <- as.data.frame(list.files(dir_iucn_shp)) %>%
  rename(shp_fn = `list.files(dir_iucn_shp)`) %>%
  filter(tools::file_ext(shp_fn) == 'shp') %>%
  mutate(shp_fn = str_replace(shp_fn, '.shp', ''))

iucn_shp_info <- data.frame()

for (spp_gp in groups_list$shp_fn) { # spp_gp <- groups_list$shp_fn[3]
  spp_dbf <- foreign::read.dbf(file.path(dir_iucn_shp, sprintf('%s.dbf', spp_gp)))
  message('file read successfully... ')
  cites <- spp_dbf %>%
    select(id_no, compiler, year, citation) %>%
    unique() %>%
    mutate(spp_group = spp_gp)

  iucn_shp_info <- bind_rows(iucn_shp_info, cites)
  message('binding to list...\n')
}

### oldest records go back to late 70s for terrestrial mammals; then skip to 2004 for a marine turtle map.

@@

***

``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = FALSE}
```

All AquaMaps species richness map:

![figures/all_aquamaps_map.png](../figures/all_aquamaps_map.png)

All IUCN species richness map:

![figures/all_iucn_map.png](../figures/all_iucn_map.png)

``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = FALSE}
```

>>>>

# Figures 1, 2

``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
```

@@

# taxa_df_long3 <- taxa_df %>%
#   dplyr::select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
#   gather(dataset, n_spp, AquaMaps:Both) %>%
#   group_by(class) %>%
#   mutate(n_total = sum(n_spp, na.rm = TRUE),
#          percent = n_spp/n_total * 100)
#
# bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
#   ggtheme_plot + 
#   geom_bar(stat = 'identity', alpha = .5) +
#   scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
#   geom_text(aes(label = sprintf('n = %s', n_total), y = 3), 
#             #size = 3, 
#             angle = 90, hjust = 0, color = 'grey30') +
#   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
#   scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
#   labs(# title = 'Species maps by taxonomic class, Red List-assessed only',
#        x = 'Class', 
#        y = 'Percent of assessed mapped species')
# print(bar_taxa3)
# ggsave(filename = file.path(dir_fig, 'si_bar_by_taxa_assessed.png'), 
#        width = 17.8, height = 6, units = 'cm')

@@

# Figure 3

## Quadrant map

### mongo plot time
scatter_quadplot <- ggplot(spp_list %>% 
                             transform(popn_category = reorder(popn_category, category_score)),
                           aes(x = area_ratio, 
                               y = sm_perc, 
                               color = popn_category,
                               size = lg_area_pct)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
#             fill= "green4")  + 
             fill= "#4dac26")  + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymin = dist_align_mean, ymax = 100, 
             alpha = .3, 
#             fill= "greenyellow") + 
             fill= "#b8e186") + 
    annotate("rect", xmin = area_align_mean, xmax = 100, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
#             fill= "orange") + 
             fill= "#f1b6da") + 
    annotate("rect", xmax = area_align_mean, xmin =   0, 
                     ymax = dist_align_mean, ymin =   0, 
             alpha = .3, 
#             fill= "orangered2") + 
             fill= "#d01c8b") + 
  geom_point(alpha = .6)

### Manage scales for color and size 
scatter_quadplot <- scatter_quadplot +
  scale_colour_manual(values = c('LC' = 'green4', 
                                 'NT' = 'greenyellow', 
                                 'VU' = 'yellow', 
                                 'EN' = 'orange', 
                                 'CR' = 'orangered2',
                                 'DD' = 'grey30')) + 
  scale_size_continuous(limits = c(0, 100),
                        breaks = c( 0,    25,    50,    75,    100), 
                        labels = c('0%', '25%', '50%', '75%', '100%')) + 
  scale_x_continuous(expand = c(0, 0), 
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%'))

### here are mean labels:
scatter_quadplot <- scatter_quadplot +
  annotate(geom = 'text',
           x = area_align_mean, y = 5,
           hjust = 0, vjust = 0,
           color = 'grey30', 
           size = 2,
           fontface = 'bold.italic', angle = 90,
           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
  annotate(geom = 'text',
           x = 5, y = dist_align_mean,
           hjust = 0, vjust = 0,
           color = 'grey30', 
           size = 2,
           fontface = 'bold.italic', angle = 0,
           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))

### here are mean lines:
# scatter_quadplot <- scatter_quadplot +
#   geom_vline(x = area_align_mean,
#              color = 'grey50', linetype = 'dashed') +
#   geom_hline(y = dist_align_mean,
#              color = 'grey50', linetype = 'dashed')

### here are individual species, identified and namean on the plot:
# scatter_quadplot <- scatter_quadplot +
#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
#   geom_text(data = spp_list_label, 
#             aes(label = sciname_label), 
#             color = 'grey30',
#             fontface = 'bold.italic',
#             size = 2,
#             hjust = .25,
#             vjust = 1.5)

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area alignment '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'), 
#       title = 'Distribution alignment vs Extent alignment',
       color = 'Risk', size = '% ocean range')

# print(scatter_quadplot)

ggsave(file.path(dir_fig, 'scatter_quadmap.png'), width = 17.8, height = 9, units = 'cm')

@@

![](../figures/scatter_quadmap.png)

## Break down quads by species group

# print(barchart_spp_gp_quads)
ggsave(plot = barchart_spp_gp_quads,
       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
       width = 17.8, height = 7, units = 'cm')

![](../figures/barchart_spp_gp_quads.png)

# Plots for SI

## Break down quadrants by risk category

# print(barchart_spp_risk_quads)
ggsave(plot = barchart_spp_risk_quads,
       filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.png'),
       width = 17.8, height = 5, units = 'cm')

```
![](../figures/si_barchart_spp_risk_quads.png)

-----

## AquaMaps threshold exploration

``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
```

-----

## Count species occurrences as proxy for data poverty

-----

# Obsolete plots; not evaluated

* bar chart: number of spp per group
* scatter plot: area compare IUCN vs AM, including log plot and plot by species group
* rangemap: Chromis alpha

``` {r child = 'data_explore_pt1.Rmd', eval = FALSE}
```

-----

* percent overlap vs total area exploration

``` {r, child = 'data_explore_pt2.Rmd', eval = FALSE}
```

-----

* Multiple species maps per quadrant

``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
```

-----

### 2016-03-21 data_explore/data_explore_am_thresh.Rmd
cleaning up data explore

```{r aquamaps_load_data, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE}

@@

```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}

@@

hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  geom_vline(xintercept = spp_df3_mean, color = 'darkred', linetype = 'dashed') + 
  annotate('text', x = spp_df3_mean, y = .16, hjust = 0, vjust = -.5,
           angle = 90, 
           #size = 2, 
           color = 'grey20',
           label = 'mean') + 
  geom_vline(xintercept = spp_df3_med, color = 'grey30', linetype = 'dashed') + 
  annotate('text', x = spp_df3_med, y = .16, hjust = 0, vjust = -.5,
           angle = 90, 
           #size = 2, 
           color = 'grey20',
           label = 'median') + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = 'Proportion of extent included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)))

print(hist_area_vs_threshold)

ggsave(file.path(dir_fig, 'AMThreshold.png'), 
       width = 17.8, height = 9, units = 'cm')
```

```{r hist_lost_area_w_threshold_ne_dd_removed, warning = FALSE, message = FALSE, echo = FALSE}

@@

hist_area_vs_threshold2 <- ggplot(spp_df3a, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  geom_vline(xintercept = spp_df3a_mean, color = 'darkred', linetype = 'dashed') + 
  annotate('text', x = spp_df3a_mean, y = .16, hjust = 0, vjust = -.5,
           angle = 90, 
           #size = 2, 
           color = 'grey20',
           label = 'mean') + 
  geom_vline(xintercept = spp_df3a_med, color = 'grey30', linetype = 'dashed') + 
  annotate('text', x = spp_df3a_med, y = .16, hjust = 0, vjust = -.5,
           angle = 90, 
           #size = 2, 
           color = 'grey20',
           label = 'median') + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = 'Proportion of extent included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)))

ggsave(file.path(dir_fig, 'AMThreshold2.png'), 
       width = 17.8, height = 9, units = 'cm')

print(hist_area_vs_threshold2)

@@

```{r scatter_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}

### create dataframe that gives me all species, their total number of cells, 
### total number of cells >= 0.4 prob, difference between the two and threat

@@

thresh_df_summary <- threshold_df %>%
  group_by(thresh) %>%
  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
            median_prop = median(proportion, na.rm = TRUE),
            quant09 = quantile(proportion, .05),
            quant25 = quantile(proportion, .25),
            quant75 = quantile(proportion, .75),
            quant91 = quantile(proportion, .95))

scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
  ggtheme_plot + 
  geom_point(aes(y = median_prop), color = 'grey60') +
  geom_point() + 
  geom_abline(slope = -1, intercept = 1, color = 'red') + 
  geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(# title = 'Mean (median) species range after presence threshold cutoff',
       x = 'Presence threshold cutoff', 
       y = 'Mean (median) relative species range')

ggsave(file.path(dir_fig, 'scatter_AMThreshold.png'), 
       width = 17.8, height = 9, units = 'cm')

print(scatter_area_vs_threshold)

@@

```{r ribbon_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}

area_mean_model <- lm(mean_prop ~ thresh, data = thresh_df_summary)
area_median_model  <- lm(median_prop ~ thresh,  data = thresh_df_summary)

@@

  geom_ribbon(aes(ymax = quant91, ymin = quant09), alpha = .5, color = NA, fill = 'grey80') +
  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
  geom_point(color = 'grey30') + 
  geom_point(aes(y = mean_prop), color = 'darkred') + 
  geom_abline(intercept = area_mean_model[['coefficients']][1],
              slope     = area_mean_model[['coefficients']][2],
              color = 'red') + 
  geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
  geom_vline(xintercept = 0.50, color = 'blue', linetype = 'dashed') +
  geom_vline(xintercept = 0.60, color = 'blue', linetype = 'dashed') +
  annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,
           angle = 90, 
           #size = 2, 
           color = 'grey20',
           label = 'Ocean Health Index: 40%') + 
  annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
           angle = 90, 
           #size = 2, 
           color = 'grey20',
           label = 'MPA Gap Analysis: 50%') + 
  annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
           angle = 90, 
           #size = 2, 
           color = 'grey20',
           label = 'Marine mammals: 60%') + 
  annotate('text', x = .75, y = .85,
           #size = 2, 
           color = 'grey20',
           label = 'R^2^ = .995') + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(# title = 'Relative species range after presence threshold',
       x = 'Presence threshold cutoff', 
       y = 'Mean (median) relative species range')

@@

ggsave(file.path(dir_fig, 'ribbon_AMThreshold.png'), 
       width = 17.8, height = 6, units = 'cm')

@@

>>>>

```{r aquamaps_load_data, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE, eval = TRUE}

@@

```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}

@@

hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  geom_vline(xintercept = spp_df3_mean, color = 'darkred', linetype = 'dashed') + 
  annotate('text', x = spp_df3_mean, y = .13, hjust = 0, vjust = -.5,
           angle = 90, 
           size = 3, 
           color = 'grey20',
           label = 'mean') + 
  geom_vline(xintercept = spp_df3_med, color = 'grey30', linetype = 'dashed') + 
  annotate('text', x = spp_df3_med, y = .13, hjust = 0, vjust = -.5,
           angle = 90, 
           size = 3, 
           color = 'grey20',
           label = 'median') + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = 'Proportion of range included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)),
       fill = 'IUCN code')

# print(hist_area_vs_threshold)

ggsave(file.path(dir_fig, 'am_threshold.png'), 
       width = 17.8, height = 9, units = 'cm')
```

![](../figures/am_threshold.png)

```{r hist_lost_area_w_threshold_ne_dd_removed, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}

@@

hist_area_vs_threshold2 <- ggplot(spp_df3a, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  geom_vline(xintercept = spp_df3a_mean, color = 'darkred', linetype = 'dashed') + 
  annotate('text', x = spp_df3a_mean, y = .16, hjust = 0, vjust = -.5,
           angle = 90, 
           size = 3, 
           color = 'grey20',
           label = 'mean') + 
  geom_vline(xintercept = spp_df3a_med, color = 'grey30', linetype = 'dashed') + 
  annotate('text', x = spp_df3a_med, y = .16, hjust = 0, vjust = -.5,
           angle = 90, 
           size = 3, 
           color = 'grey20',
           label = 'median') + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = 'Proportion of range included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)),
       fill = 'IUCN code')

ggsave(file.path(dir_fig, 'am_threshold2.png'), 
       width = 17.8, height = 9, units = 'cm')

@@

![](../figures/am_threshold2.png)

```{r df_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}
### create dataframe that gives me all species, their total number of cells, 
### total number of cells >= 0.4 prob, difference between the two and threat

@@

thresh_df_summary <- threshold_df %>%
  group_by(thresh) %>%
  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
            median_prop = median(proportion, na.rm = TRUE),
            quant05 = quantile(proportion, .05),
            quant25 = quantile(proportion, .25),
            quant75 = quantile(proportion, .75),
            quant95 = quantile(proportion, .95))

@@

```{r ribbon_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE, eval = TRUE}

@@

  geom_ribbon(aes(ymax = quant95, ymin = quant05), alpha = .5, color = NA, fill = 'grey80') +
  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
  geom_line(color = 'grey30') + 
  # geom_line(aes(y = mean_prop), color = 'darkgreen') + 
  # geom_abline(intercept = area_mean_model[['coefficients']][1],
  #             slope     = area_mean_model[['coefficients']][2],
  #             color = 'red') + 
  geom_vline(xintercept = 0.40, color = 'blue', linetype = 'dashed') +
  geom_vline(xintercept = 0.50, color = 'blue', linetype = 'dashed') +
  geom_vline(xintercept = 0.60, color = 'blue', linetype = 'dashed') +
  annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,
           angle = 90, 
           size = 3, 
           color = 'grey20',
           label = 'Halpern et al. 2012') + 
  annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
           angle = 90, 
           size = 3, 
           color = 'grey20',
           label = 'Klein et al. 2015') + 
  annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
           angle = 90, 
           size = 3, 
           color = 'grey20',
           label = 'Kaschner et al. 2006') + 
  # annotate('text', x = .75, y = .85,
  #          #size = 3, 
  #          color = 'grey20',
  #          label = 'R^2^ = .995') + 
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(  .2,    .4,    .5,    .6,    .8,   1.0),
                     labels = c('20%', '40%', '50%', '60%', '80%', '100%')
                     limits = c(0, 1.02)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(# title = 'Relative species range after presence threshold',
       x = 'AquaMaps presence threshold cutoff', 
       y = 'Relative species range')

@@

ggsave(file.path(dir_fig, 'ribbon_am_threshold.png'), 
       width = 17.8, height = 6, units = 'cm')

@@

![](../figures/ribbon_am_threshold.png)

### 2016-03-21 data_explore/data_explore_global_maps.Rmd
cleaning up data explore

```{r setup, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

@@

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           progress = TRUE,
                           col_types = 'cddddc')

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), 
                           progress = TRUE)

### LOICZID cell ID numbers - used in plotting rasters
loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'
```

```{r global_maps, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

library(rasterVis)

myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))

message('Full AM dataset, species count\n')

all_am_rast_file   <- file.path(dir_data, 'rasters/am_all_rast.tif')

dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')

if(!file.exists(all_am_rast_file)) {
  am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
    as.data.frame(stringsAsFactors = FALSE)

@@

png(file.path(dir_fig, 'all_aquamaps_map1.png'),
    width = 8.7, height = 6, units = 'in', res = 300)
print(am_levelplot1)
dev.off()

@@

Now look at full IUCN species

```{r all_iucn, eval = FALSE, echo = FALSE}

message('Full IUCN dataset, species count\n')

@@

if(!file.exists(all_iucn_rast_file)) {
  iucn_cells_file    <- file.path(dir_data_iucn, 'iucn_cells_2015.csv')
  iucn_spp_cells_all <- fread(iucn_cells_file) %>%
        as.data.frame(stringsAsFactors = FALSE)

@@

} else {
  all_iucn_map   <- raster(all_iucn_rast_file)
  values(all_iucn_map)[is.na(values(all_am_map))] <- NA
}

@@

png(file.path(dir_fig, 'all_iucn_map1.png'),
    width = 8.7, height = 6, units = 'in', res = 300)
print(iucn_levelplot1)
dev.off()

@@

``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
### rescale values relative to maximum cell value

am_map_max <- maxValue(all_am_map)
all_am_pct_map <- all_am_map/am_map_max

iucn_map_max <- maxValue(all_iucn_map)
all_iucn_pct_map <- all_iucn_map/iucn_map_max

diff_pct_map    <- all_am_pct_map - all_iucn_pct_map

### plot maps with distributions as percentages.  Should it be 0 - 1, or 
### 0 - max value?
levelplot(all_am_pct_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset percent of max')
levelplot(all_iucn_pct_map, par.settings = myTheme, 
          main = 'IUCN Dataset percent of max')
### this is the difference one. Should it be min value to max value?
levelplot(diff_pct_map, par.settings = myTheme, 
          main = 'Difference: AquaMaps - IUCN by percent')

diff_ln_map <- all_am_ln_map - all_iucn_ln_map
levelplot(diff_ln_map, 
          par.settings = myTheme, 
          pretty = TRUE,
          margin = list(axis = TRUE),
#          main = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', 
          colorkey = list(at = seq(-1, 1, length = 15)))
maxValue(all_am_ln_map)
minValue(all_am_ln_map)
maxValue(all_iucn_ln_map)
minValue(all_iucn_ln_map)
maxValue(diff_ln_map)
minValue(diff_ln_map)
diff_ln_pts <- as.data.frame(rasterToPoints(diff_ln_map))

spp_richness_diff_plot <- ggplot(diff_ln_pts, aes(x = x, y = y)) +
  ggtheme_map + 
  geom_raster(aes(fill = layer)) +
  scale_fill_gradient2(low = 'red', mid = 'white', high = 'blue', midpoint = 0) +
  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
  labs(title = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', x = NULL, y = NULL)

spp_richness_diff_plot

>>>>

```{r setup_from_data_explore, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

@@

### All AquaMaps species count by cell

```{r all_am_map, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

library(rasterVis)
myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))

all_am_rast_file   <- file.path(dir_data, 'rasters/am_all_rast.tif')
dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')

if(!file.exists(all_am_rast_file)) {
  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

  am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
    as.data.frame(stringsAsFactors = FALSE)

@@

png(file.path(dir_fig, 'all_aquamaps_map.png'),
    width = 8.7, height = 6, units = 'in', res = 300)
print(am_levelplot1)
dev.off()

@@

![](../figures/all_aquamaps_map.png)

### All IUCN species count by cell

```{r all_iucn, eval = TRUE, echo = FALSE}

@@

if(!file.exists(all_iucn_rast_file)) {
  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

  iucn_cells_file    <- file.path(dir_data_iucn, 'iucn_cells_2015.csv')
  iucn_spp_cells_all <- fread(iucn_cells_file) %>%
        as.data.frame(stringsAsFactors = FALSE)

@@

} else {
  all_iucn_map   <- raster(all_iucn_rast_file)
}

@@

png(file.path(dir_fig, 'all_iucn_map.png'),
    width = 8.7, height = 6, units = 'in', res = 300)
print(iucn_levelplot1)
dev.off()

@@

![](../figures/all_iucn_map.png)
